Evolution of pre-copulatory and post-copulatory inbreeding behaviour

### 2016-09-27 ms_pre_post.Rmd
Tweak and add new figures

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding avoidance has been widely observed through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no conceptual framework compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: Will evolution of post-copulatory inbreeding avoidance strengthen selection for polyandry? Will selection for one costly inbreeding strategy be negligible given a non-costly complementary strategy? And will selection for one inbreeding strategy be negligible given existence of an already adaptive alternative strategy? We show that evolution of polyandry is greatly facilitated by post-copulatory inbreeding avoidance. Strong selection for relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

@@

<!--- It is therefore necessary to consider evolution of inbreeding mating strategy and inbreeding fertilisation strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation. --->

<!--- Overall, to predict evolution of polyandry and inbreeding strategies, it is therefore necessary to model costs associated with polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy. --->

<!--- Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent allele copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [Duthie and Reid *in press*; @Duthie]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'inbreeding mating strategy') and post-copulatory inbreeding strategy (hereafter 'inbreeding fertilisation strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.  --->

<!--- Critically, evolution of inbreeding fertilisation strategy is precluded if females only mate with one male, meaning that evolution of polyandry (defined as female multiple mating within a single reproductive bout) is intrinsically tied to female inbreeding fertilisation strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015].  However, no models quantify the degree to which expression of biased paternity resulting from evolving inbreeding fertilisation strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness. --->

<!--- Evolution of inbreeding fertilisation strategy might increase selection for polyandry, but evolutionary feedbacks between inbreeding fertilisation strategy and polyandry might also be contingent upon evolution of inbreeding mating strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and inbreeding fertilisation strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their inbreeding mating strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on inbreeding fertilisation strategy might occur to reinforce inbreeding mating strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of inbreeding mating strategy and inbreeding fertilisation strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive. --->

<!--- Evolution of polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy will ultimately be determined by the, potentially interacting, fitness benefits and costs of each phenotype. Given severe inbreeding depression, phenotypes will increase a focal female's fitness if they cause her to avoid breeding with close relatives [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However,  --->

<!--- Selection for inbreeding mating strategy versus inbreeding fertilisation strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for inbreeding mating strategy and weakening selection for polyandry. Alternatively, evolution of female inbreeding mating strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through inbreeding fertilisation strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). --->

@@

Here we use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance causes selection for evolution of polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that selection for pre-copulatory inbreeding avoidance but not post-copulatory inbreeding avoidance will occur when pre-copulatory inbreeding avoidance is less costly than the cost of post-copulatory inbreeding avoidance, while selection for post-copulatory inbreeding avoidance but not pre-copulatory inbreeding avoidance will occur when pre-copulatory inbreeding avoidance is more costly than the cost of post-copulatory inbreeding avoidance. Finally, we test the hypothesis that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.

<!--- that constraints on evolution of one inbreeding strategy (e.g., inbreeding mating strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., inbreeding fertilisation strategy), and that --->

@@

We track interactions among discrete individuals in a small focal population that persists over multiple generations (key individual traits, parameter values, and variables are described in Table 1). Individuals within this population are either female or male, and can be related to each other if they share a common ancestor. We record the coefficient of kinship ($k$), defined as the probability that a randomly selected homologous allele from each individual is identical-by-descent [@Lynch1998], between all pairs of individuals in each generation using the population pedigree [see @Boyce1983; @Duthie]. At the start of each generation, each female pays up to three independent costs of (1) tendency to engage in polyandry ($c_{P}$), (2) mating strategy with respect to inbreeding ($c_{M}$), and (3) fertilisation strategy with respect to inbreeding ($c_{F}$). The number of males that a focal female mates with ($N_{males}$) is determined by the effects of 20 alleles ($P_{a}$) across 10 diploid loci that combine additively to determine her genotype ($P_{g}$), and ultimately her phenotype for tendency to engage in polandry ($P_{p}$). A focal female's choice of mate(s) is affected by 20 alleles ($M_{a}$) across 10 diploid loci that combine additively to affect her genotype ($M_{g}$) and phenotype ($M_{p}$) for pre-copulatory inbreeding avoidance or preference, and kinship between the focal female and her potential mates. Similarly, after mating has occurred, post-copulatory fertilisation of a focal female's $n$ total offspring is affected by 20 alleles ($F_{a}$) across 10 diploid loci that combine additively to affect the genotype ($F_{g}$) and phenotype ($F_{p}$) of the degree to which fertilisation success is affected by kinship. After fertilisation, the viability of a focal female's $n$ offspring is decreased by inbreeding depression ($\beta$), and surviving offspring become adults. After inbreeding depression in offspring viability, $\rho$ adult male immigrants arrive in the population, then density effects reduce the entire population down to separate female ($K_{f}$) and male ($K_{m}$) carrying capacities. Over generations, we record mean values of alleles underlying polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

@@

<!--- In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For inbreeding mating strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for inbreeding fertilisation strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. --->

Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), pre-copulatory inbreeding strategy (i.e., inbreeding mating strategy, $M_{p}$), and post-copulatory inbreeding strategy (i.e., inbreeding fertilisation strategy, $F_{p}$) each have costs. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of inbreeding mating strategy and inbreeding fertilisation strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a function of the coefficient of kinship between $i$ and the sire of her offspring $j$ ($k_{ij}$), and a log-linear slope of inbreeding depression $\beta$. We model selection as having an absolute effect on offspring viability (i.e., hard selection), rather than a relative effect on viability (i.e., soft selection), so that the effect of $\beta$ is consistent over generations and across different parameter combinations. Natural selection on tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$) can therefore be compared across generations and parameter combinations. If offspring viability was instead relative such that the least inbred offspring had a high viability regardless of the absolute value of $k_{ij}$, then the effect of kinship coefficients and inbreeding depression slopes on offspring viability would depend on the population-wide distribution of $k_{ij}$ between females and males, making model inference more difficult. Therefore, we define offspring viability as,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
The slope $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might combine to become homozygous in offspring and reduce viability. Equation 1 assumes that each deleterious recessive allele affects offspring viability independently, meaning that every deleterious recessive allele added will have a multiplicative affect on decreasing total viability [@Morton1956; @Mills1994], causing the log-linear relationship between parent kinship and inbreeding depression. We assume that inbreeding depression always decreases offspring viability, $\beta > 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, the product of $-\beta$ and $k_{ij}$ must be a value less than or equal to zero, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We can therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability using a Bernoulli trial.

@@

After offspring viability is realised, $\rho$ adult immigrants are added to the focal population. The coefficient of kinship between an immigrant and all other individuals (including native individuals and other immigrants) always equals zero ($k_{ij}=0$). Because kinship coefficient values are calculated directly from the pedigree of the focal population, the addition of immigrants thereby introduces non-kin into the population at a constant rate and prevents the mean kinship coefficient of the population from asymptoting to a value of 1 over generations. If the focal population was instead closed ($\rho=0$), then all alleles would quickly be expected to become identical-by-descent among descendants of the ancestral population. To avoid having immigrants affect evolution of tendency for polyandry ($P_{p}$), inbreeding mating strategy ($M_{p}$), and inbreeding fertilisation strategy ($F_{p}$), immigrants are always male and can therefore be chosen based on their kinship values of zero but not actively affect reproductive decisions through the expression of $P_{p}$, $M_{p}$, or $F_{p}$. To further avoid having immigrants affect evolutionary dynamics, allele values underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) are randomly sampled from normal distributions with means and standard deviations of $P_{a}$, $M_{a}$, and $F_{a}$ values equal to the means and standard deviations of those in the focal population at the time of immigration. Overall, this model of immigration effectively assumes that the focal population receives immigrants from other nearby populations that are subject to the same selection on $P_{p}$, $M_{p}$, and $F_{p}$.

@@

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), inbreeding mating strategy ($M_{a}$), and inbreeding fertilisation strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different parameter combinations to test our three focal hypotheses.

To test our hypothesis that selection for polyandry is driven by a female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve given $c_{P} = \{0, 0.0025, 0.005,  0.01\}$. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship (i.e, $F_{p}=0$). Hence, while mutation and drift can still occur for $F_{a}$ alleles, these alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance. By comparing mean $P_{a}$ values over generations when post-copulatory inbreeding avoidance also evolves versus when fertilisation is random, we test whether or not evolution of post-copulatory inbreeding strategy increases selection for polyandry.

<!--- We simulate up to eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), inbreeding mating strategy ($c_{M}$), and inbreeding fertilisation strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be a strong cost of tendency for polyandry] for values of $\beta = 3$ ($8$ total combinations). --->

@@

Individual traits (A), model parameter values (B), and model variables (C) for an individual based model of the evolution of polyandry, inbreeding mating strategy, and inbreeding fertilisation strategy.

>>>>

---
title: "Evolution of pre-copulatory and post-copulatory inbreeding behaviour"
author: "A. Bradley Duthie et al."
date: "August 24, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding avoidance has been widely observed through pre-copulatory and post-copulatory mechanisms as a consequence of inbreeding depression in offspring viability. However, no conceptual framework compares evolution of pre-copulatory and post-copulatory inbreeding avoidance jointly, and given biologically relevant costs of inbreeding strategies. Critically, evolution of post-copulatory inbreeding avoidance is restricted to females that mate multiply, potentially leading to complex evolutionary dynamics among inbreeding strategies and polyandry. Accordingly, we use individual-based modelling to track evolution of polyandry and inbreeding strategies under multiple costs and constraints to address three questions: Will evolution of post-copulatory inbreeding avoidance strengthen selection for polyandry? Will selection for a costly inbreeding strategy be negligible given a non-costly complementary strategy? And will selection for an inbreeding strategy be negligible given existence of an already adaptive alternative strategy? We show that selection for polyandry is greatly strengthened by post-copulatory inbreeding avoidance. Strong selection for relatively low cost inbreeding avoidance strategies was observed, while selection for functionally redundant high cost strategies decreased over generations. Finally, existence of an adaptive inbreeding strategy precluded evolution of post-copulatory, but not pre-copulatory, inbreeding avoidance. Our model thereby introduces a novel framework for predicting evolution of both pre-copulatory and post-copulatory inbreeding strategies in empirical systems.

@@

<!--- It is therefore necessary to consider evolution of pre-copulatory inbreeding strategy and post-copulatory inbreeding strategy when each evolves in a population where its complementary inbreeding strategy is evolutionarily constrained or already fixed as an adaptation. --->

<!--- Overall, to predict evolution of polyandry and inbreeding strategies, it is therefore necessary to model costs associated with polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy. --->

<!--- Inbreeding, defined as breeding between relatives, directly affects inheritence of identical-by-descent alleles and the viability of inbred offspring, meaning that the decision to prefer, avoid, or tolerate (i.e., mate randomly with respect to relatedness) inbreeding is predicted to have profound effects on individual fitness [@Parker2006; @Kokko2006; @Szulkin2012; @Duthie2015a]. Viability of inbred offspring typically decreases with increasing relatedness between parents [termed 'inbreeding depression'; @Keller2002; @Charlesworth2009]. If inbreeding depression is sufficiently weak, selection for female inbreeding preference is predicted because inbreeding increases the reproductive success of a focal female's male relative, thereby increasing the number of identical-by-descent allele copies present in her offspring [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However, if inbreeding depression is moderate to strong, selection for female inbreeding avoidance is predicted [Duthie and Reid *in press*; @Duthie]. Female inbreeding strategy can be enacted before or after copulation occurs [e.g., @Tuni2013; @Firman2015; @Daniel2015; @Fischer2015]. However, models of inbreeding do not consider pre-copulatory inbreeding strategy (hereafter 'pre-copulatory inbreeding strategy') and post-copulatory inbreeding strategy (hereafter 'post-copulatory inbreeding strategy') jointly, meaning that it is unclear when selection for either or both strategies is predicted in empirical systems.  --->

<!--- Critically, evolution of post-copulatory inbreeding strategy is precluded if females only mate with one male, meaning that evolution of polyandry (defined as female multiple mating within a single reproductive bout) is intrinsically tied to female post-copulatory inbreeding strategy. Although polyandry is commonplace in wild populations, the regularity with which it is observed in the wild remains theoretically puzzling because engaging in polyandry does not necessarily increase the number of offspring that a female produces, meaning that the fitness benefits associated with polyandry are not always clear [@Parker2013; @Taylor2014]. It has been hypothesised that females can acheive a net fitness benefit if polyandry provides them with additional opportunities to adjust their inbreeding through copulations with additional males [e.g., @Tregenza2002; @Akcay2007; @Lehtonen2015].  However, no models quantify the degree to which expression of biased paternity resulting from evolving post-copulatory inbreeding strategy can feedback to drive evolution of polyandry in populations of potentially inbreeding females and males with biologically realistic distributions of relatedness. --->

<!--- Evolution of post-copulatory inbreeding strategy might increase selection for polyandry, but evolutionary feedbacks between post-copulatory inbreeding strategy and polyandry might also be contingent upon evolution of pre-copulatory inbreeding strategy, thereby leading to complex evolutionary dynamics between inbreeding strategies and polyandry. For example, if females can acheive their desired degree of inbreeding through mate choice alone, then subsequent polyandry and post-copulatory inbreeding strategy might be superfluous for increasing female fitness. Alternatively, if females are often unsuccessful at realising their pre-copulatory inbreeding strategy due to factors such as sexual conflict over mating and resulting male coercion [e.g., @Parker2006], limited kin recognition amongst potential mates [e.g., @Penn2010], or simply the inherent stochasticity associated with mate encounters or expression of mate choice based on inbreeding [e.g., @Kokko2006; @Duthie], then selection on post-copulatory inbreeding strategy might occur to reinforce pre-copulatory inbreeding strategy. In empirical systems where inbreeding depression is severe and inbreeding avoidance is therefore beneficial, expression of pre-copulatory inbreeding strategy and post-copulatory inbreeding strategy have been observed across diverse taxa [e.g., @Gasparini2011; @Tuni2013; @Firman2015; @Fischer2015; @Daniel2015]. However, no models predict *a priori* which of these inbreeding strategies should be observed when inbreeding depression is sufficiently severe to make inbreeding avoidance adaptive. --->

<!--- Evolution of polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy will ultimately be determined by the, potentially interacting, fitness benefits and costs of each phenotype. Given severe inbreeding depression, phenotypes will increase a focal female's fitness if they cause her to avoid breeding with close relatives [@Parker1979; @Parker2006; @Kokko2006; @Duthie2015a]. However,  --->

<!--- Selection for pre-copulatory inbreeding strategy versus post-copulatory inbreeding strategy and associated polyandry might also depend evolutionary constraints, or on existing phenotypes in the population. For example, if phenotypes biasing fertilisation to avoid inbreeding are precluded by evolutionary constraints [@Bradshaw1991; @Futuyma2010], then inbreeding avoidance might only be possible before copulation occurs, perhaps strengthening selection for pre-copulatory inbreeding strategy and weakening selection for polyandry. Alternatively, evolution of female pre-copulatory inbreeding strategy might be precluded due to severely limited kin recognition or the inability to resist male coercion given sexual conflict [@Arnqvist2005a; @Parker2006; @Penn2010]. In such conditions, females might only be able to express choice through post-copulatory inbreeding strategy, perhaps strengthening selection for inbreeding avoidance through polyandry and post-copulatory mechanisms. Selection for inbreeding mating or fertilisation strategy might also be weak or negligible if the strategy is redundant due to the presence of an already established adaptive inbreeding strategy (e.g., if females already successfully avoid inbreeding during mate encounters, then they might be unable to increase their fitness further by biasing fertilisation after mating has occurred). --->

@@

Here we use individual-based modelling and *in silico* experiments to test the hypothesis that, given severe inbreeding depression in offspring viability, evolution of post-copulatory inbreeding avoidance causes selection for polyandry. Additionally, we test the hypothesis that costs associated with polyandry and inbreeding strategies will strongly affect evolution of each such that selection for pre-copulatory inbreeding avoidance but not post-copulatory inbreeding avoidance will occur when pre-copulatory inbreeding avoidance is less costly than the cost of post-copulatory inbreeding avoidance, while selection for post-copulatory inbreeding avoidance but not pre-copulatory inbreeding avoidance will occur when pre-copulatory inbreeding avoidance is more costly than the cost of post-copulatory inbreeding avoidance. Finally, we test the hypothesis that evolution of one inbreeding strategy will be precluded if it is introduced into a population that is already fixed for a complementary inbreeding strategy that is adaptive. We thereby provide a framework for predicting evolution of pre-copulatory and post-copulatory inbreeding avoidance across reproductive systems with different relative costs and strong inbreeding depression in offspring viability.

<!--- that constraints on evolution of one inbreeding strategy (e.g., pre-copulatory inbreeding strategy) will increase the strength of selection on evolution of a complementary inbreeding strategy (e.g., post-copulatory inbreeding strategy), and that --->

@@

We track interactions among discrete individuals in a small focal population that persists over multiple generations (key individual traits, parameter values, and variables are described in Table 1). Individuals within this population are either female or male, and can be related to each other if they share a common ancestor. We record the coefficient of kinship ($k$), defined as the probability that a randomly selected homologous allele from each individual is identical-by-descent [@Lynch1998], between all pairs of individuals in each generation using the population pedigree [see @Boyce1983; @Duthie]. At the start of each generation, each female pays up to three independent costs of (1) tendency to engage in polyandry ($c_{P}$), (2) mating strategy with respect to inbreeding ($c_{M}$), and (3) fertilisation strategy with respect to inbreeding ($c_{F}$). The number of males that a focal female mates with ($N_{males}$) is determined by the effects of 20 alleles ($P_{a}$) across 10 diploid loci that combine additively to determine her genotype ($P_{g}$), and ultimately her phenotype for tendency to engage in polandry ($P_{p}$). A focal female's choice of mate(s) is affected by 20 alleles ($M_{a}$) across 10 diploid loci that combine additively to affect her genotype ($M_{g}$) and phenotype ($M_{p}$) for pre-copulatory inbreeding avoidance or preference, and kinship between the focal female and her potential mates. Similarly, after mating has occurred, post-copulatory fertilisation of a focal female's $n$ total offspring is affected by 20 alleles ($F_{a}$) across 10 diploid loci that combine additively to affect the genotype ($F_{g}$) and phenotype ($F_{p}$) of the degree to which fertilisation success is affected by kinship. After fertilisation, the viability of a focal female's $n$ offspring is decreased by inbreeding depression ($\beta$), and surviving offspring become adults. After inbreeding depression in offspring viability, $\rho$ adult male immigrants arrive in the population, then density effects reduce the entire population down to separate female ($K_{f}$) and male ($K_{m}$) carrying capacities. Over generations, we record mean values of alleles underlying polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy.

@@

<!--- In the case of tendency to engage in polyandry, the cost of $P_{p}$ ($c_{P}$) is interpreted biologically as a cost associated with mate searching or courtship [e.g., @Rowe1988; @Rowe1994; @Ronkainen1994; @Koga1998; @Lasley-Rasher2012]. For pre-copulatory inbreeding strategy, the cost of $M_{p}$ ($c_{M}$) is interpreted biologically as a cost associated with risk of mating failure due to females being overly choosy [e.g., @Kokko2013], and for post-copulatory inbreeding strategy, the cost of $F_{p}$ ($c_{F}$) is interpreted as an energetic cost associated with the physiological or biochemical mechanisms that need to be developed to bias fertilisation for inbreeding avoidance or inbreeding preference [e.g., @Gasparini2011]. --->

Values of phenotypic traits for tendency to engage in polyandry ($P_{p}$), pre-copulatory inbreeding strategy (i.e., inbreeding mating strategy, $M_{p}$), and post-copulatory inbreeding strategy (i.e., inbreeding fertilisation strategy, $F_{p}$) each have costs. Each cost independently increases the risk of female pre-mating mortality and, because generations are non-overlapping, total reproductive failure as a linear function of its associated trait. The probability that a focal female dies before mating due to the cost of polyandry is therefore $P_{p} \times c_{P}$. Similarly, the probability of pre-mating female mortality due to the cost of pre-copulatory inbreeding strategy and post-copulatory inbreeding strategy is $|M_{p}| \times c_{M}$ and $|F_{p}| \times c_{F}$, repectively, where $|M_{p}|$ is the absolute value of $M_{p}$ and $|F_{p}|$ is the absolute value of $F_{p}$. Absolute values are used for applying costs to inbreeding strategies because both negative and positive $M_{p}$ and $F_{p}$ bias breeding interactions with respect to kinship, while only positive values of $P_{p}$ affect a female's tendency for polyandry; zero and negative $P_{p}$ values always translate to monandrous phenotypes (see below).

@@

The viability of a focal female $i$'s offspring ($\Psi_{\textrm{off}}$) decreases as a function of the coefficient of kinship between $i$ and the sire of her offspring $j$ ($k_{ij}$), and a log-linear slope of inbreeding depression $\beta$. We model selection as having an absolute effect on offspring viability (i.e., hard selection), rather than a relative effect on viability (i.e., soft selection), so that the effect of $\beta$ is consistent over generations and across different parameter combinations. Natural selection on tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$) can therefore be compared across generations and parameter combinations. If offspring viability was instead relative such that the least inbred offspring had a high viability regardless of the absolute value of $k_{ij}$, then the effect of kinship coefficients and inbreeding depression slopes on offspring viability would depend on the population-wide distribution of $k_{ij}$ between females and males, making model inference more difficult. Therefore, we define offspring viability as,
\begin{equation}
\Psi_{\textrm{off}} = e^{-\beta k_{ij}}
\end{equation}
The slope $\beta$ models the number of haploid lethal equivalents that exist as deleterious recessive alleles in the gametes of $i$ and $j$, and which might combine to become homozygous in offspring and reduce viability. Equation 1 assumes that each deleterious recessive allele affects offspring viability independently, meaning that every deleterious recessive allele added will have a multiplicative affect on decreasing total viability [@Morton1956; @Mills1994], causing the log-linear relationship between parent kinship and inbreeding depression. We assume that inbreeding depression always decreases offspring viability, $\beta > 0$ (i.e., no outbreeding depression). Therefore, because $k_{ij}$ is constrained to values between zero and one, the product of $-\beta$ and $k_{ij}$ must be a value less than or equal to zero, $-\beta \times k_{ij} \leq 0$. Values of $\Psi_{\textrm{off}}$ must therefore be between zero (if $-\beta \times k_{ij}$ is very negative) and one (if $-\beta \times k_{ij} = 0$). We can therefore define $\Psi_{\textrm{off}}$ as the probability that an offspring is viable, and sample its realised viability using a Bernoulli trial.

@@

After offspring viability is realised, $\rho$ adult immigrants are added to the focal population. The coefficient of kinship between an immigrant and all other individuals (including native individuals and other immigrants) always equals zero ($k_{ij}=0$). Because kinship coefficient values are calculated directly from the pedigree of the focal population, the addition of immigrants thereby introduces non-kin into the population at a constant rate and prevents the mean kinship coefficient of the population from asymptoting to a value of 1 over generations. If the focal population was instead closed ($\rho=0$), then all alleles would quickly be expected to become identical-by-descent among descendants of the ancestral population. To avoid having immigrants affect evolution of tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory inbreeding strategy ($F_{p}$), immigrants are always male and can therefore be chosen based on their kinship values of zero but not actively affect reproductive decisions through the expression of $P_{p}$, $M_{p}$, or $F_{p}$. To further avoid having immigrants affect evolutionary dynamics, allele values underlying tendency for polyandry ($P_{a}$), pre-copulatory inbreeding strategy ($M_{a}$), and post-copulatory inbreeding strategy ($F_{a}$) are randomly sampled from normal distributions with means and standard deviations of $P_{a}$, $M_{a}$, and $F_{a}$ values equal to the means and standard deviations of those in the focal population at the time of immigration. Overall, this model of immigration effectively assumes that the focal population receives immigrants from other nearby populations that are subject to the same selection on $P_{p}$, $M_{p}$, and $F_{p}$.

@@

To analyse our model, we track mean values of alleles underlying tendency for polyandry ($P_{a}$), pre-copulatory inbreed strategy ($M_{a}$), and post-copulatory inbreeding strategy ($F_{a}$) in a focal population. We present evolutionary dynamics of these mean allele values across different parameter combinations to test our three focal hypotheses.

To test our hypothesis that selection for polyandry is driven by a female's ability to bias fertilisation toward inbreeding avoidance through post-copulatory mechanisms, we compare simulations in which all three phenotypes including tendency for polyandry ($P_{p}$), pre-copulatory inbreeding strategy ($M_{p}$), and post-copulatory strategy ($F_{p}$) evolve, with an identical set of simulations in which $F_{p}$ cannot evolve given $c_{P} = \{0, 0.0025, 0.005,  0.01\}$. To stop $F_{p}$ from evolving, we sever the connection from $F_{a}$ to $F_{p}$ such that all $F_{g}$ genotypes resulting from $F_{a}$ cause random fertilisation with respect to kinship (i.e, $F_{p}=0$). Hence $F_{a}$ alleles have no phenotypic effect and polyandry cannot cause polyandrous females to benefit through an increased ability to bias fertilisation toward inbreeding avoidance. By comparing mean $P_{a}$ values over generations when post-copulatory inbreeding avoidance also evolves versus when fertilisation is random, we test whether or not evolution of post-copulatory inbreeding strategy increases selection for polyandry.

<!--- We simulate up to eight possible permutations of values of costs of tendency for polyandry ($c_{P}$), pre-copulatory inbreeding strategy ($c_{M}$), and post-copulatory inbreeding strategy ($c_{F}$) equal to either $0$ or $0.02$ [@Duthie observed $0.02$ to be a strong cost of tendency for polyandry] for values of $\beta = 3$ ($8$ total combinations). --->

@@

Individual traits (A), model parameter values (B), and model variables (C) for an individual based model of the evolution of polyandry, pre-copulatory inbreeding strategy, and post-copulatory inbreeding strategy.
